---
title: "Midpoint Deliverable"
author: "Keith Ellingwood, Sage Choi, Audrey Kho, Faduma Edey"
date: "2/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("knitr")
source("scripts/summary_calc.R")
source("scripts/summary_table.R")
source("scripts/first_chart.R")
source("scripts/second_chart.R")
source("scripts/third_chart.R")

data <- read.csv("data/pollution_us_2000_2016.csv")
```

## The Problem
For our project, we wanted to explore the field of environmental trends. All of our group members are concerned about environmental trends that we are seeing, especially because climate change seems to already have reached such an advanced stage while people in authority mostly continue to remain unbothered by it. We wanted to use data to learn more about it while also educating others about what exactly is happening with our planet's environment. By answering some of the questions we were curious about, we hope to contribute to the solution to this extremely urgent problem. One particular issue we were all concerned about is the effects of climate change, and what factors may be contributing to it.

```{r, echo = FALSE}
summary_list <- get_summary_info(data)
```

## Summary Information
The dataset we are using contains information on **pollution in the United States** from the years **2000-2016**. The data comes from _kaggle.com_, and was sourced from the United States Environmental Protection Agency _(US EPA)_ database. There are **`r summary_list["columns"]`** columns and **`r summary_list["rows"]`** rows in the data. The data contains lots of information about the locations in which the readings were taken, and includes things like **`r unlist(summary_list["column_names"])`**. There are **`r summary_list["city_count"]`** different cities in the US accounted for in this dataset. It focuses on 4 main types of greenhouse gasses: **`r unlist(summary_list["gasses"])`**.

## Table
Here, we have a table of our data aggregated by State and sorted from highest total emmissions to lowest. Each state contains the amount of each major pollutant we are looking at total for the years 2000-2016, as well as all of the pollutatnts totaled (in parts per billion notation). We included this so we could see a breakdown of exactly which states in the US are the biggest culprits for high emmissions, and to also see which pollutants were higher in which states. From the chart, we can see that **California** has the highest total emissions for all pollutants in the US, clocking in at **281,350,906.5** PPB total. Contrary to that, we can see that **Washington** actually has the lowest total emmissions, at **227,664.4** PPB total. This is suprising because we expected **Rhode Island** to be the lowest since it is the state with the smallest amount of landmass.

```{r, echo = FALSE}
kable(get_summary_table(data), col.names = names_col)
```

## Charts
<!-- Three different charts that display information from the data (what you
visualize is up to you). Describe the purpose of each chart (e.g., "this chart
was intended to show the relationship between..."), render each chart, and
describe pertinent observations found in each chart. --> 
### Chart 1
The line chart below shows the **average yearly emissions levels** of each of the four air pollutants in the dataset: _Carbon Monoxide (CO)_, _Sulfur Dioxide (SO2)_, _Ozone (O3)_, and _Nitrogen Dioxide (NO2)_. The data shown in this chart summarizes the emission levels of these harmful gases from the years 2000-2016 and is intended to show the gases' emission levels relative to each other. More importantly, this chart shows the changes in the average emission levels of each gas through the years.
</p>
As can be seen in the chart, the average _CO_ emissions have consistently been the highest by a couple hundred parts per billion throughout the years. The average _O3_, _NO2_, and _SO2_ emission levels have been quite close to each other all throughout and saw very slight changes. On the other hand, the line representing the _CO_ emissions has clearly been experiencing a downward trend since the year 2000, which is great news for the environment. However, a slight rise can also be observed towards the end of the _CO_ emission graph, which could be an indicator of a shift into an upward trend in the following years.

```{r fig.height = 6, fig.width = 10.5, echo= FALSE}
chart_1 <- print(line_chart(data) + labs(y = "Average gas emission (ppb)", x = "Year") +
          ggtitle("Average emission levels of O3, CO, NO2, and SO2 gases from 2000-2016") +
            theme(plot.title = element_text(size = 16, face = "bold")))
```

### Chart 2
Chart 2 is a pie chart intended to show the relationship between the averages of **four major pollutants**: Carbon Monoxide **(CO)**, Sulfur Dioxide **(SO2)**, Ozone **(O3)**, and Nitrogen Dioxide **(NO2)** in part per billion. 
</p>
When observing the chart below, three observations come into view. The first is that Carbon Monoxide has the **largest mean**. By knowing which major pollutant has the highest average out of 1.75 million observations, researchers can focus on what sources and reasons are causing such a high number. A second observation is that averages give researchers an easier time comparing how the pollutants stack up with each other. By doing this, these pollutants can possibly be prioritized based on their rank.

```{r, echo = FALSE}
chart_2 <- pie_chart(data)
```

### Chart 3
Chart 3 is a mapping of all the total emissions from 2000 to 2016. It is actually the same data used for our table above, but mapped onto a chloropleth map to show each state's differences comparatively. This map makes it clearly visual that **California's** total emissions outpace nearly every other state by a huge factor. What's interesting is that there are a few states that don't show up on the map, **Montana, Mississippi, Nebraska and West Virginia**. Looking at the table we made earlier, there is no entries for these states either.

```{r, echo = FALSE}
geo_chart(data)
```