---
title: "Midpoint Deliverable"
author: "Keith Ellingwood, Sage Choi, Audrey Kho, Faduma"
date: "2/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("knitr")
source("scripts/summary_calc.R")
source("scripts/summary_table.R")
source("scripts/first_chart.R")
source("scripts/second_chart.R")
source("scripts/third_chart.R")

data <- read.csv("data/pollution_us_2000_2016.csv")
```

## The Problem
For our project, we wanted to explore the field of environmental trends. All of our group members are concerned about environmental trends that we are seeing, especially because climate change seems to already have reached such an advanced stage while people in authority mostly continue to remain unbothered by it. We wanted to use data to learn more about it while also educating others about what exactly is happening with our planet's environment. By answering some of the questions we were curious about, we hope to contribute to the solution to this extremely urgent problem. One particular issue we were all concerned about is the effects of climate change, and what factors may be contributing to it.

```{r, echo = FALSE}
summary_list <- get_summary_info(data)
```
## Summary Information
Our dataset we are using is information on pollution in the United States from the years 2000-2016. The data comes from kaggle.com, and was sourced from the Harvard Database and hosted on the WorldFish Dataverse under Harvard Dataverse. There are `r summary_list["columns"]` columns and `r summary_list["rows"]` rows in the data. The data contains lots of information about the locations in which the readings were taken, and includes things like the . There are  different cities in the US accounted for in this dataset. It focuses on 4 main types of greenhouse gasses: .

## Table
<!-- A table of aggregated data (e.g., that has been "grouped" by one of your
features) with pertinent observations found in it. # You must also describe why
you included the table, and what information it reveals.-->


## Charts
<!-- Three different charts that display information from the data (what you
visualize is up to you). Describe the purpose of each chart (e.g., "this chart
was intended to show the relationship between..."), render each chart, and
describe pertinent observations found in each chart. --> 