---
title: "Midpoint Deliverable"
author: "Keith Ellingwood, Sage Choi, Audrey Kho, Faduma"
date: "2/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("knitr")
source("scripts/summary_calc.R")
source("scripts/summary_table.R")
source("scripts/first_chart.R")
source("scripts/second_chart.R")
source("scripts/third_chart.R")

data <- read.csv("data/pollution_us_2000_2016.csv")
```

## The Problem
For our project, we wanted to explore the field of environmental trends. All of our group members are concerned about environmental trends that we are seeing, especially because climate change seems to already have reached such an advanced stage while people in authority mostly continue to remain unbothered by it. We wanted to use data to learn more about it while also educating others about what exactly is happening with our planet's environment. By answering some of the questions we were curious about, we hope to contribute to the solution to this extremely urgent problem. One particular issue we were all concerned about is the effects of climate change, and what factors may be contributing to it.

```{r, echo = FALSE}
summary_list <- get_summary_info(data)
```

## Summary Information
Our dataset we are using is information on **pollution in the United States** from the years **2000-2016**. The data comes from _kaggle.com_, and was sourced from the Harvard Database and hosted on the WorldFish Dataverse under Harvard Dataverse. There are **`r summary_list["columns"]`** columns and **`r summary_list["rows"]`** rows in the data. The data contains lots of information about the locations in which the readings were taken, and includes things like the date, county, city, etc. There are  different cities in the US accounted for in this dataset. It focuses on 4 main types of greenhouse gasses: Carbon Monoxide, Sulfur Dioxide, Ozone, and Nitrogen Dioxide.

## Table
<!-- A table of aggregated data (e.g., that has been "grouped" by one of your
features) with pertinent observations found in it. # You must also describe why
you included the table, and what information it reveals.-->


## Charts
<!-- Three different charts that display information from the data (what you
visualize is up to you). Describe the purpose of each chart (e.g., "this chart
was intended to show the relationship between..."), render each chart, and
describe pertinent observations found in each chart. --> 
### Chart 1
The line chart below shows the **average yearly emissions levels** of each of the four air pollutants in the dataset: _Carbon Monoxide (CO)_, _Sulfur Dioxide (SO2)_, _Ozone (O3)_, and _Nitrogen Dioxide (NO2)_. The data shown in this chart summarizes the emission levels of these harmful gases from the years 2000-2016 and is intended to show the gases' emission levels relative to each other. More importantly, this chart shows the changes in the average emission levels of each gas through the years.
</p>
As can be seen in the chart, the average _CO_ emissions have consistently been the highest by a couple hundred parts per billion throughout the years. The average _O3_, _NO2_, and _SO2_ emission levels have been quite close to each other all throughout and saw very slight changes, if any. On the other hand, the line representing the _CO_ emissions has clearly been experiencing a downward trend since the year 2000, which is great news for the environment. However, a slight rise can also be observed towards the end of the _CO_ emission graph, which could be an indicator of a shift into an upward trend in the following years.

```{r fig.height = 6, fig.width = 10.5, echo= FALSE}
chart_1 <- print(line_chart(data) + labs(y = "Average gas emission (ppb)", x = "Year") +
          ggtitle("Average emission levels of O3, CO, NO2, and SO2 gases from 2000-2016"))
```

### Chart 2
Chart 2 is a pie chart intended to show the relationship between the averages of **four major pollutants**: Carbon Monoxide **(CO)**, Sulfur Dioxide **(SO2)**, Ozone **(O3)**, and Nitrogen Dioxide **(NO2)** in part per billion. 
</p>
When observing the chart below, three observations come into view. The first is that Carbon Monoxide has the **largest mean**. By knowing which major pollutant has the highest average out of 1.75 million observations, researchers can focus on what sources and reasons are causing such a high number. A second observation is that averages give researchers an easier time comparing how the pollutants stack up with each other. By doing this, these pollutants can possibly be prioritized based on their rank.

```{r, echo = FALSE}
chart_2 <- pie_chart(data)
```